---
title: "R Notebook"
output: html_notebook
Developer: Yogesh Awdhut Gadade
---

```{r}
acfma2=ARMAacf(ma=c(0.5,0.4), lag.max=10) # 10 lags of ACF for MA(2) with theta1 = 0.5 theta2 = 0.4
acfma2
```

# The first command determines the ACF and stores it in an object named acfma2 (our choice of name). 
# Printed numerical values of the ACF simply use the command acfma2.

```{r}
lags=0:10                       #creates a variable named lags that ranges from 0 to 10.
plot(lags,acfma2,xlim=c(1,10), ylab="r",type="h", main = "ACF for MA(2) with theta1 = 0.5,theta2=0.4")
abline(h=0)                     #adds a horizontal axis to the plot
```
# The above plot command plots lags versus the ACF values for lags 1 to 10. 
# The ylab parameter labels the y-axis and the "main" parameter puts a title on the plot.

#The simulation and plots were done with the following commands:
```{r}
xc=arima.sim(n=150, list(ma=c(0.5, 0.4)))
x=xc+10               # adds 10 to make mean = 10. Simulation defaults to mean = 0.
plot(x, type="b", main = "Simulated MA(2) Series")
```
#simulated n = 150 values from this model and plotted the sample time series and the sample ACF for the simulated data. 

```{r}
acf(x, xlim=c(1,10), main="ACF for simulated MA(2) Data")
```






