---
title: "R Notebook"
output: html_notebook
Developer: Yogesh Awdhut Gadade
Date: 21st Feb 2022
---
# Q1 a)
```{r}
acfma2=ARMAacf(ma=c(0.5,0.4), lag.max=10) # 10 lags of ACF for MA(2) with theta1 = 0.5 theta2 = 0.4
acfma2
```

The first command determines the ACF and stores it in an object named acfma2 (our choice of name). 
Printed numerical values of the ACF simply use the command acfma2.

```{r}
lags=0:10                       #creates a variable named lags that ranges from 0 to 10.
plot(lags,acfma2,xlim=c(1,10), ylab="r",type="h", main = "ACF for MA(2) with theta1 = 0.5,theta2=0.4")
abline(h=0)                     #adds a horizontal axis to the plot
```
The above plot command plots lags versus the ACF values for lags 1 to 10. 
The ylab parameter labels the y-axis and the "main" parameter puts a title on the plot.

The simulation and plots were done with the following commands:
```{r}
xc=arima.sim(n=150, list(ma=c(0.5, 0.4)))
x=xc+10               # adds 10 to make mean = 10. Simulation defaults to mean = 0.
plot(x, type="b", main = "Simulated MA(2) Series")
```
simulated n = 150 values from this model and plotted the sample time series and the sample ACF for the simulated data. 

```{r}
acf(x, xlim=c(1,10), main="ACF for simulated MA(2) Data")
```



# Q1 b)
```{r}
acfma2=ARMAacf(ma=c(1.2,-0.7), lag.max=10) # 10 lags of ACF for MA(2) with theta1 = 0.5 theta2 = 0.4
acfma2
```

The first command determines the ACF and stores it in an object named acfma2 (our choice of name). 
Printed numerical values of the ACF simply use the command acfma2.

```{r}
lags=0:10                       #creates a variable named lags that ranges from 0 to 10.
plot(lags,acfma2,xlim=c(1,10), ylab="r",type="h", main = "ACF for MA(2) with theta1 = 1.2, theta2=-0.7")
abline(h=0)                     #adds a horizontal axis to the plot
```
The above plot command plots lags versus the ACF values for lags 1 to 10. 
The ylab parameter labels the y-axis and the "main" parameter puts a title on the plot.

The simulation and plots were done with the following commands:
```{r}
xc=arima.sim(n=150, list(ma=c(1.2,-0.7)))
x=xc+10               # adds 10 to make mean = 10. Simulation defaults to mean = 0.
plot(x, type="b", main = "Simulated MA(2) Series")
```
simulated n = 150 values from this model and plotted the sample time series and the sample ACF for the simulated data. 

```{r}
acf(x, xlim=c(1,10), main="ACF for simulated MA(2) Data")
```




# ############# Method 2 
# Q1 a)
```{r}
par(mfrow=c(3,1))
acfma1 <- ARMAacf(ma=c(0.5,0.4),lag.max=10)
lags <- 0:10
plot(lags,acfma1,type="h",main="ACF for MA(2) with theta1=0.5 and theta2=0.4")
acfma2 <- ARMAacf(ma=c(1.2,-.7),lag.max=10)
plot(lags,acfma2,type="h",main="ACF for MA(2) with theta1=1.2 and theta2=-0.7")
```

# Q1 a)
The simulation and related R program is given below:
```{r}
par(mfrow=c(3,3))
acfma1 <- ARMAacf(ma=c(0.5,0.4),lag.max=10)
lags <- 0:10
plot(lags,acfma1,type="h")
t1 <- arima.sim(n = 100, list(ma = c(0.5, 0.4)))
t2 <- arima.sim(n = 100, list(ma = c(0.5, 0.4)))
acf(t1,lag.max=10)
acf(t2,lag.max=10)
acfma2 <- ARMAacf(ma=c(1.2,-.7),lag.max=10)
plot(lags,acfma2,type="h")
t1 <- arima.sim(n = 100, list(ma = c(1.2, -0.7)))
t2 <- arima.sim(n = 100, list(ma = c(1.2, -0.7)))
acf(t1,lag.max=10)
acf(t2,lag.max=10)
acfma3 <- ARMAacf(ma=c(-1,-.6),lag.max=10)
plot(lags,acfma2,type="h")
```





