---
title: "R Notebook"
output: html_notebook
Author: Yogesh Awdhut Gadade
---
### Q4 Sketch the autocorrelation functions for each of the following ARMA models:
### (a) ARMA(1,1), with phi = 0.7 and theta = 0.4
### (b) ARMA(1,1) with phi = 0.7 and theta = -0.4
### Hint: Like question 1., you can use ARMAacf function.

### (c) Simulate each the models in (a)-(b) with 150 observations, and plot theirs series, and their sample ACF.

### Q4 a)  phi = 0.7 and theta = 0.4

```{r}
acfma2=ARMAacf(ar=0.7, ma=0.4, lag.max=10) # 10 lags of ACF for ARMA(1, 1) with phi = 0.7 and theta = 0.4
acfma2
```
```{r}
lags=0:10                       #creates a variable named lags that ranges from 0 to 10.
plot(lags,acfma2,xlim=c(1,10), ylab="r",type="h", main = "ACF for ARMA(1, 1) with phi = 0.7 and theta = 0.4")
abline(h=0)                     #adds a horizontal axis to the plot
```
### Q4 (b) ARMA(1,1) with phi = 0.7 and theta = -0.4
```{r}
acfma2=ARMAacf(ar=0.7, ma=-0.4, lag.max=10) # 10 lags of ACF for ARMA(1, 1) with phi = 0.7 and theta = -0.4
acfma2
```
```{r}
lags=0:10                       #creates a variable named lags that ranges from 0 to 10.
plot(lags,acfma2,xlim=c(1,10), ylab="r",type="h", main = "ACF for ARMA(1, 1) with phi = 0.7 and theta = -0.4")
abline(h=0)                     #adds a horizontal axis to the plot
```
### (c) Simulate each the models in (a)-(b) with 150 observations, and plot theirs series, and their sample ACF.
```{r}
### a) phi = 0.7 and theta = 0.4
xc=arima.sim(n=150, list(ma=c(0.7,0.4)))
x=xc+10               # adds 10 to make mean = 10. Simulation defaults to mean = 0.
plot(x, type="b", main = "Simulated ARMA(1, 1) Series with phi = 0.7 and theta = 0.4")
```
```{r}
acf(x, xlim=c(1,10), main="ACF for simulated ARMA(1, 1) Data Series with phi = 0.7 and theta = 0.4")

```
#### Observation: As it can be seen here that some of the bars are going out side the confidance area hence it is non-stationary


```{r}
### b) phi = 0.7 and theta = -0.4
xc=arima.sim(n=150, list(ma=c(0.7,-0.4)))
x=xc+10               # adds 10 to make mean = 10. Simulation defaults to mean = 0.
plot(x, type="b", main = "Simulated ARMA(1, 1) Series with phi = 0.7 and theta = -0.4")
```
```{r}
acf(x, xlim=c(1,10), main="ACF for simulated ARMA(1, 1) Data Series with phi = 0.7 and theta = -0.4")

```
#### Observation: It can be seen that some of the bars are outside the confidance area hence we can say that its non-stationary




#### Exploring other methods: ( Can be removed - experimenting )


# ------------------------ Q4 Method 1 --------------------------

### Q4
### Sketch the autocorrelation functions for each of the following ARMA models:
### (a) ARMA(1,1) with phi = 0.7 and theta = 0.4.
### (b) ARMA(1,1) with phi = 0.7 and theta = -0.4.
### (a) ARMA(1,1) with phi = 0.7 and theta = 0.4. Solution:

### Q4 (a) ARMA(1,1) with phi = 0.7 and theta = 0.4.
```{r}
a1 <- list();        # created a1 list
a1 <- append(a1,0)
a1
```

```{r}
noise <- rnorm(5000) #simulates random variate having a specified normal distribution.
noise[1:10]
```

```{r}
for (i in 2:500){
  a1 <- append(a1, (a1[[i-1]]*.7 + noise[i]) + (noise[i] - .4*noise[i-1]));
}
a1 <- unlist(a1)   # unlist function used to convert a list to vector
plot(a1, type="o", col="orange")
```

```{r}
acf(a1)
```

### Q4 (b) ARMA(1,1) with phi = 0.7 and theta = −0.4. Solution:Similarly

```{r}
a2 <- list();
a2 <- append(a2,0)
for (i in 2:500){
  a2 <- append(a2,(a2[[i-1]]*.7 + noise[i]) + (noise[i] - (-.4*noise[i-1])));
}
a2 <- unlist(a2)
plot(a2, type="o", col="orange")
```

```{r}
acf(a2)
```

# ------------------------ Q4 Method 2 --------------------------

###################################################### 

### Time Series Analysis
### Yogesh Awdhut Gadade
### Objective - Investigate ARMA models

###################################################### 

```{r}
rm(list=ls())
options(width=200)
# install if not present
#install.packages("astsa")
```

```{r}
#library(astsa)
cat(" Investigate ARMA models\n\n", "Last modification:",date(),'\n')
```

###################################################### 

###################################################### 

# Sketch the autocorrelation functions for each of the following ARMA models:

# (a) ARMA(1,1) with φ = 0.7 and θ = 0.4.

# (b) ARMA(1,1) with φ = 0.7 and θ = −0.4.

# Hint: Like question 1., you can use ARMAacf function. (c) Simulate each the

# models in (a)-(b) with 150 observations, and plot theirs series, and their sample ACF.

```{r}
#  (a) ARMA(1,1) with φ = 0.7 and θ = 0.4.
ARMA11 <- arima.sim(model=list(ar=.7,ma=.4),n=150)
#ARMA01 <- arima.sim(model=list(ma=.7),n=150)
#ARMA10 <- arima.sim(model=list(ar=.4),n=150)
plot(ARMA11, type="o", col="orange")
```

```{r}
#
# compute ACF for 
#
ARMA11.acf <- acf(ARMA11,plot=F)
#ARMA01.acf <- acf(ARMA01,plot=F)
#ARMA10.acf <- acf(ARMA10,plot=F)
#
# compute theoretical ACFs for all models
#
acf.11 <- ARMAacf(ar=.7,ma=.4,lag.max=20)     ## Example from Brockwell & Davis
#acf.01 <- ARMAacf(ma=.7,lag.max=20)
#acf.10 <- ARMAacf(ar=.4,lag.max=20)
#
# plot against each other
#
plot(ARMA11.acf)
#lines(ARMA11.acf$lag,acf.11)
```

```{r}
#  (b) ARMA(1,1) with φ = 0.7 and θ = -0.4.
ARMA11 <- arima.sim(model=list(ar=.7,ma=-.4),n=150)
plot(ARMA11, type="o", col="orange")
```

```{r}
# compute ACF for 
ARMA11.acf <- acf(ARMA11,plot=F)
# compute theoretical ACFs for all models
acf.11 <- ARMAacf(ar=.7,ma=-.4,lag.max=20)
# plot against each other
plot(ARMA11.acf)
```
































