---
title: "R Notebook"
output: html_notebook
---
# Yogesh Awdhut Gadade
# Date: 11th April 2022
# Robot Data Analysis

#The data file named robot contains a time series obtained from an industrial robot.
#The robot was put through a sequence of maneuvers, and the distance from a
#desired ending point was recorded in inches. This was repeated 324 times to form
#the time series.
```{r}
library(TSA)
rm(list=ls(all=TRUE))
data(robot)
```
```{r}
plot(robot, main="Time Series Plot of the industrial robot data")
```
# Comment: The time series plot appears to be stationary.
```{r}
summary(robot)
```
# Sample ACF, PACF and EACF
```{r}
acf(robot)
```
```{r}
pacf(robot)
```
```{r}
eacf(robot)
```
# Comment: The ACF and PACF plots shows that the time series is
# stationary. We can confirm it with adf test.
```{r}
#Augmented Dickey-Fuller (ADF) Test in R
#H0: The time series is non-stationary. In other words, it has some time-dependent structure 
#and does not have constant variance over time.
#HA: The time series is stationary.
library(tseries)
adf.test(robot) # stationary because p value is less than 0.05
```
```{r}
## ARMA Subset
SUB=armasubsets(robot,nar=5,nma=5)
plot(SUB)
```
# Comment: The BIC is minimum for ARMA(1,1) model as seen in the Plot.

# (a) Estimate the parameters of an AR(1) model for these data.
```{r}
# Fitting AR(1)
arima(robot, order = c(1,0,0))
```
#(b) Estimate the parameters of an IMA(1,1) model for these data.
```{r}
# Fitting ARIMA
arima(robot, order = c(0,1,1))
```
#(c) Compare the results from parts (a) and (b) in terms of AIC.

#As it can be seen that the values of AIC for AR(1) is -2947.08 and for IMA(1, 1) is -2959.9 
#which is slightly different and more in case of IMA(1, 1)


















