---
title: "R Notebook"
output: html_notebook
---

# Quarterly U.S. GNP series, Y, from 1941(first quarter) to 1991(first quarter), n =
# 177 is given in a Excel, text, and SAS data set called “gnp”. This data set is online
# in HW folder, and in a folder called Data.

```{r}
library(astsa)   # There are lots of fun data sets included in astsa
#Seasonally adjusted quarterly U.S. GNP from 1947(1) to 2002(3).
#help(astsa)
#data(package = "astsa")
```

The data frame econ5 was used to consider the effect of quarterly GNP, 
consumption, and government and private investment on U.S. unemployment. 
```{r}
gr = diff(log(econ5))
gr = scale(gr)  # for comparable spectra
tsplot(gr, ncol=2, col=2:6, lwd=2, byrow=FALSE)
```
In this case, mvspec will plot the individual spectra by default:
```{r}
gr.spec = mvspec(gr, spans=c(7,7), detrend=FALSE, taper=.25, col=2:6, lwd=2, main='spectra')
```
You can extract the spectral matrices as fxx, an array of dimensions dim = c(p,p,nfreq) 
as well as plot coherencies and phases. Here, p = 5:
```{r}
round(gr.spec$fxx, 2)
```
```{r}
round(var(econ5), 1)
```
```{r}
gr  # As it can be seen econ5 consists of gnp column
```
```{r}
data(gnp)# in astsa pckage or library we have econ5 data consists of 5 columns one of it is gnp
plot(gnp) 
```
```{r}
gnp
```
#  ---------------------------- Using data from 1947 to 1991

In our case we have been provided excel file consists of Quarterly U.S. GNP series, Y, from 1941(first quarter) to 1991(first quarter), n =
177 is given in a Excel, text, and SAS data set called “gnp”. This data set is online
in HW folder, and in a folder called Data. Answer the following questions:
```{r}
#install.packages("readxl")
library("readxl")

Y <- read_excel("gnp2.xls")
Y
```

```{r}
# Plotting gnp from astsa lobrary data eco5 gnp from 1947 to 2002
plot(gnp)
```
# a) Give a plot of the original series.
```{r}
#Plotting value of gnp from given input excel 1947 to 1991
plot(x = Y$DATE,y = Y$gnp,
     xlab = "x-axis",
     ylab = "y-axis",
     main = "Plot",
     type='l'
)
```


#b) Determine if the data needs variance stabilization and give the transformation
#the Box-Cox procedure suggests. 

```{r}
library(MASS)

boxcox(lm(Y$gnp ~ 1))
```
Note that the center dashed vertical line represents the estimated parameter \hat{\lambda} 
and the others the 95% confidence interval of the estimation.
As the plot shows that the 0 is inside the confidence interval of the optimal 
\lambda and as the estimation of the parameter is really close to 0 in this example, 
the best option is to apply the logarithmic transformation of the data 

#Box-cox tranformation: fitting gnp to y
```{r}
boxcox(Y$gnp ~ Y$y)
```


#c) Transform the GNP series, Y , to log(Y).
```{r}
Y$Loggnp = log(Y$gnp)
Y
```
```{r}
plot(x = Y$DATE,y = Y$Loggnp, xlab = "x-axis", ylab = "y-axis", main = "Plot", type='l')
```
#d) Now, address the trend in the Box-Cox transformed series and the log(Y ) by taking 
# simple difference of each series. Comment on the ACF of each of the differenced
# series.
#Box-cox tranformation: fitting gnp to y
```{r}
boxcox(Y$gnp ~ Y$y)
```
```{r}
#fit linear regression model
model <- lm(Y$gnp ~ Y$y)
#find optimal lambda for Box-Cox transformation 
bc <- boxcox(Y$gnp ~ Y$y)
(lambda <- bc$x[which.max(bc$y)])
model
```

```{r}
#fit new linear regression model using the Box-Cox transformation
new_model <- lm(((Y$y^lambda-1)/lambda) ~ Y$gnp)
new_model
```
#Taking difference of Log series:
```{r}
diff(Y$Loggnp)
```
#Plotting box-cox graph for the differenced log series:
```{r}
boxcox(lm(diff(Y$gnp) ~ 1))
```
Since some of the values are negative hence can not draw box-cox graph. Now let's
see the ACF graph for the gnp:
```{r}
acf(Y$gnp)
```
# ACF graph for the Log gnp:
```{r}
acf(Y$Loggnp)
```
# ACF graph for the differenced Log gnp:
```{r}
acf(diff(Y$Loggnp))
```
Which is much stable than previous two graphs.

Now check out the ACF for the box cox transformed series
```{r}
acf(new_model$residuals, main="ACF for Box Cox transformed series")
```
# ACF of Differenced box-cox transformed series:
```{r}
acf(diff(new_model$residuals), main="ACF for differenced Box Cox transformed series")
```
compared to differenced log series this one seems less stable.


#e) Plot the histogram (overlay the appropriate normal distribution) and Q-Q plot
#of each of the transformed & simple-differenced series. Comment on these.
```{r}
#define plotting area
op <- par(pty = "s", mfrow = c(1, 2))
#Now checking the histograms of box-cox tranformed series:
hist(new_model$residuals)
# Histogram of differenced box-cox transformed series:
hist(diff(new_model$residuals))
#display both histograms
par(op)
```
```{r}
#define plotting area
op <- par(pty = "s", mfrow = c(1, 2))
# Histogram for Log transformed GNP
hist(Y$Loggnp)
# Histogram of differenced Log transformed GNP
hist(diff(Y$Loggnp))

#display both Histograms
par(op)
```
```{r}
#define plotting area
op <- par(pty = "s", mfrow = c(1, 2))

#Q-Q plot for original model
qqnorm(model$residuals)
qqline(model$residuals)

#Q-Q plot for Box-Cox transformed model
qqnorm(new_model$residuals)
qqline(new_model$residuals)

#display both Q-Q plots
par(op)
```
# Propose at least three good ARIMA models and discuss (i) parameter estimation, (ii) AIC and SBC (BIC) numbers, 
# (iii) residual analysis (e.g., Q-Q plot, Ljung-Box-Pierce Test) and the whole
# To check MA side let's draw PACF as well along with ACF of difference log data
```{r}
#define plotting area
op <- par(pty = "s", mfrow = c(1, 2))
# Histogram for Log transformed GNP
acf(diff(Y$Loggnp))
# Histogram of differenced Log transformed GNP
pacf(diff(Y$Loggnp))
#display both Histograms
par(op)
```
